---
title: "Tasks for attendees"
format: html
editor_options: 
  chunk_output_type: console
---

## Task 1: Create a flow map visualizing total flight trips originating from NYC in December

```{r nyc-flows-dec}
# write your code here
```

Compare and contrast NYC flow maps for May and December.

::: callout-tip
## Point to ponder

Wouldn't it be easier if we could see month-by-month flow changes in a single view?
:::

## Task 2: Create a flow map visualizing hourly fluctuations in flight trips originating from NYC on Christmas eve

```{r nyc-flows-christmas-eve}
# write your code here
```

## Task 3: Create a flow map visualizing hourly fluctuations in flight trips originating from NYC on Christmas

```{r}
# write your code here
```

Compare and contrast NYC flow maps for Christmas Eve and Christmas.

::: callout-tip
## Point to ponder

Wouldn't it be easier if we could see flows for Christmas Eve and Christmas in a single view?
:::

## Task 4: Create a flow map comparing hourly fluctuations in flight trips originating from NYC on Christmas Eve vs. Christmas
